---
title: "Table S3 to S10. 2019 opioid-related mortality by area"
author: "Mathew Kiang"
date: "5/19/2021"
output: 
  html_document: 
    code_folding: 'hide'
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, results='hide'}
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)
library(DT)
source(here("code", "utils.R"))
```

```{r}
results_df <-
    readr::read_csv(here::here("data", "joinpoint_estimates_suppressed.csv"))

results_df <- results_df %>%
    categorize_opioids() %>%
    categorize_race()

wide_df <- readRDS(here("data", "ratios_and_diffs_suppressed.RDS")) %>%
    filter(year == 2019) 
```

```{r}
summarized_table <- results_df %>%
    filter(year == 2019) %>%
    transmute(
        abbrev,
        name,
        opioid_cat,
        opioid_type,
        race_eth,
        race_cat,
        deaths = ifelse(is.na(deaths), "<11", formatC(deaths, big.mark = ",")),
        obs_rate = sprintf("%0.1f", round(obs_rate, 1)),
        model_rate = sprintf("%0.1f", round(model_rate, 1)),
        apc_print = sprintf(
            "%i (%i, %i)",
            round(apc),
            round(apc_lower),
            round(apc_upper)
        ),
        apc_pval = case_when(apc_pval < .001 ~ "<.001",
                             apc_pval < .01 ~ "<.01", 
                             TRUE ~ sprintf("%0.2f", round(apc_pval, 2))),
        doubling_time = sprintf("%0.1f (%0.1f, %0.1f)",
                                round(1 / log(1 + apc/100), 1),
                                round(1 / log(1 + apc_lower/100), 1),
                                round(1 / log(1 + apc_upper/100), 1)
        )
    ) 

write_csv(summarized_table, here("output", "tables_S3_to_S10_data_printable.csv"))
```

## Table S3. Opioid-related mortality in 2019 by state and race/ethnicity
```{r}
summarized_table %>%
    filter(nchar(abbrev) == 2,
           opioid_type == "opioid") %>%
    arrange(name) %>% 
    select(-abbrev, -opioid_cat, -opioid_type, -race_cat) %>%
    pivot_wider(
        id_cols = name,
        names_from = race_eth,
        values_from = c(deaths, obs_rate, model_rate, apc_print, apc_pval)
    ) %>%
    janitor::clean_names() %>%
    select(name,
           deaths_nhb,
           obs_rate_nhb,
           model_rate_nhb,
           apc_print_nhb,
           apc_pval_nhb,
           deaths_nhw,
           obs_rate_nhw,
           model_rate_nhw,
           apc_print_nhw,
           apc_pval_nhw) %>%
    kable(booktabs = TRUE,
          col.names = c("State", rep(
              c("Deaths", "Observed rate", "Modeled rate", "APC (95% CI)", "APC P-value"), 2
          )),
          format.args = list(big.mark = ",")) %>%
    kable_classic() %>% 
    add_header_above(c(
        "",
        "Non-Hispanic Black" = 5,
        "Non-Hispanic White" = 5
    ))
```

## Table S4. Opioid-related mortality in 2019 by city/county and race/ethnicity
```{r}
summarized_table %>%
    filter(nchar(abbrev) > 2,
           opioid_type == "opioid") %>%
    arrange(abbrev) %>% 
    select(-name, -opioid_cat, -opioid_type, -race_cat) %>%
    pivot_wider(
        id_cols = abbrev,
        names_from = race_eth,
        values_from = c(deaths, obs_rate, model_rate, apc_print, apc_pval)
    ) %>%
    janitor::clean_names() %>%
    select(abbrev,
           deaths_nhb,
           obs_rate_nhb,
           model_rate_nhb,
           apc_print_nhb,
           apc_pval_nhb,
           deaths_nhw,
           obs_rate_nhw,
           model_rate_nhw,
           apc_print_nhw,
           apc_pval_nhw) %>%
    kable(booktabs = TRUE,
          col.names = c("Area", rep(
              c("Deaths", "Observed rate", "Modeled rate", "APC (95% CI)", "APC P-value"), 2
          )),
          format.args = list(big.mark = ",")) %>%
    kable_classic() %>% 
    add_header_above(c(
        "",
        "Non-Hispanic Black" = 5,
        "Non-Hispanic White" = 5
    ))
```

## Table S5. Heroin-related mortality in 2019 by state and race/ethnicity
```{r}
summarized_table %>%
    filter(nchar(abbrev) == 2,
           opioid_type == "heroin") %>%
    arrange(name) %>% 
    select(-abbrev, -opioid_cat, -opioid_type, -race_cat) %>%
    pivot_wider(
        id_cols = name,
        names_from = race_eth,
        values_from = c(deaths, obs_rate, model_rate, apc_print, apc_pval)
    ) %>%
    janitor::clean_names() %>%
    select(name,
           deaths_nhb,
           obs_rate_nhb,
           model_rate_nhb,
           apc_print_nhb,
           apc_pval_nhb,
           deaths_nhw,
           obs_rate_nhw,
           model_rate_nhw,
           apc_print_nhw,
           apc_pval_nhw) %>%
    kable(booktabs = TRUE,
          col.names = c("State", rep(
              c("Deaths", "Observed rate", "Modeled rate", "APC (95% CI)", "APC P-value"), 2
          )),
          format.args = list(big.mark = ",")) %>%
    kable_classic() %>% 
    add_header_above(c(
        "",
        "Non-Hispanic Black" = 5,
        "Non-Hispanic White" = 5
    ))
```

## Table S6. Heroin-related mortality in 2019 by city/county and race/ethnicity
```{r}
summarized_table %>%
    filter(nchar(abbrev) > 2,
           opioid_type == "opioid") %>%
    arrange(abbrev) %>% 
    select(-name, -opioid_cat, -opioid_type, -race_cat) %>%
    pivot_wider(
        id_cols = abbrev,
        names_from = race_eth,
        values_from = c(deaths, obs_rate, model_rate, apc_print, apc_pval)
    ) %>%
    janitor::clean_names() %>%
    select(abbrev,
           deaths_nhb,
           obs_rate_nhb,
           model_rate_nhb,
           apc_print_nhb,
           apc_pval_nhb,
           deaths_nhw,
           obs_rate_nhw,
           model_rate_nhw,
           apc_print_nhw,
           apc_pval_nhw) %>%
    kable(booktabs = TRUE,
          col.names = c("Area", rep(
              c("Deaths", "Observed rate", "Modeled rate", "APC (95% CI)", "APC P-value"), 2
          )),
          format.args = list(big.mark = ",")) %>%
    kable_classic() %>% 
    add_header_above(c(
        "",
        "Non-Hispanic Black" = 5,
        "Non-Hispanic White" = 5
    ))
```

## Table S7. Natural/semi-synthetic-related mortality in 2019 by state and race/ethnicity
```{r}
summarized_table %>%
    filter(nchar(abbrev) == 2,
           opioid_type == "natural") %>%
    arrange(name) %>% 
    select(-abbrev, -opioid_cat, -opioid_type, -race_cat) %>%
    pivot_wider(
        id_cols = name,
        names_from = race_eth,
        values_from = c(deaths, obs_rate, model_rate, apc_print, apc_pval)
    ) %>%
    janitor::clean_names() %>%
    select(name,
           deaths_nhb,
           obs_rate_nhb,
           model_rate_nhb,
           apc_print_nhb,
           apc_pval_nhb,
           deaths_nhw,
           obs_rate_nhw,
           model_rate_nhw,
           apc_print_nhw,
           apc_pval_nhw) %>%
    kable(booktabs = TRUE,
          col.names = c("State", rep(
              c("Deaths", "Observed rate", "Modeled rate", "APC (95% CI)", "APC P-value"), 2
          )),
          format.args = list(big.mark = ",")) %>%
    kable_classic() %>% 
    add_header_above(c(
        "",
        "Non-Hispanic Black" = 5,
        "Non-Hispanic White" = 5
    ))
```

## Table S8. Natural/semi-synthetic-related mortality in 2019 by city/county and race/ethnicity
```{r}
summarized_table %>%
    filter(nchar(abbrev) > 2,
           opioid_type == "opioid") %>%
    arrange(abbrev) %>% 
    select(-name, -opioid_cat, -opioid_type, -race_cat) %>%
    pivot_wider(
        id_cols = abbrev,
        names_from = race_eth,
        values_from = c(deaths, obs_rate, model_rate, apc_print, apc_pval)
    ) %>%
    janitor::clean_names() %>%
    select(abbrev,
           deaths_nhb,
           obs_rate_nhb,
           model_rate_nhb,
           apc_print_nhb,
           apc_pval_nhb,
           deaths_nhw,
           obs_rate_nhw,
           model_rate_nhw,
           apc_print_nhw,
           apc_pval_nhw) %>%
    kable(booktabs = TRUE,
          col.names = c("Area", rep(
              c("Deaths", "Observed rate", "Modeled rate", "APC (95% CI)", "APC P-value"), 2
          )),
          format.args = list(big.mark = ",")) %>%
    kable_classic() %>% 
    add_header_above(c(
        "",
        "Non-Hispanic Black" = 5,
        "Non-Hispanic White" = 5
    ))
```

## Table S9. Synthetic opioid-related mortality in 2019 by state and race/ethnicity
```{r}
summarized_table %>%
    filter(nchar(abbrev) == 2,
           opioid_type == "synth") %>%
    arrange(name) %>% 
    select(-abbrev, -opioid_cat, -opioid_type, -race_cat) %>%
    pivot_wider(
        id_cols = name,
        names_from = race_eth,
        values_from = c(deaths, obs_rate, model_rate, apc_print, apc_pval)
    ) %>%
    janitor::clean_names() %>%
    select(name,
           deaths_nhb,
           obs_rate_nhb,
           model_rate_nhb,
           apc_print_nhb,
           apc_pval_nhb,
           deaths_nhw,
           obs_rate_nhw,
           model_rate_nhw,
           apc_print_nhw,
           apc_pval_nhw) %>%
    kable(booktabs = TRUE,
          col.names = c("State", rep(
              c("Deaths", "Observed rate", "Modeled rate", "APC (95% CI)", "APC P-value"), 2
          )),
          format.args = list(big.mark = ",")) %>%
    kable_classic() %>% 
    add_header_above(c(
        "",
        "Non-Hispanic Black" = 5,
        "Non-Hispanic White" = 5
    ))
```

## Table S10. Synthetic opioid-related mortality in 2019 by city/county and race/ethnicity
```{r}
summarized_table %>%
    filter(nchar(abbrev) > 2,
           opioid_type == "synth") %>%
    arrange(abbrev) %>% 
    select(-name, -opioid_cat, -opioid_type, -race_cat) %>%
    pivot_wider(
        id_cols = abbrev,
        names_from = race_eth,
        values_from = c(deaths, obs_rate, model_rate, apc_print, apc_pval)
    ) %>%
    janitor::clean_names() %>%
    select(abbrev,
           deaths_nhb,
           obs_rate_nhb,
           model_rate_nhb,
           apc_print_nhb,
           apc_pval_nhb,
           deaths_nhw,
           obs_rate_nhw,
           model_rate_nhw,
           apc_print_nhw,
           apc_pval_nhw) %>%
    kable(booktabs = TRUE,
          col.names = c("Area", rep(
              c("Deaths", "Observed rate", "Modeled rate", "APC (95% CI)", "APC P-value"), 2
          )),
          format.args = list(big.mark = ",")) %>%
    kable_classic() %>% 
    add_header_above(c(
        "",
        "Non-Hispanic Black" = 5,
        "Non-Hispanic White" = 5
    ))
```

## Table S11. 2019 rates, rate differences, and rate ratios by state
```{r}
wide_df %>%
  filter(opioid_type == "opioid", nchar(abbrev) == 2) %>%
  arrange(name) %>%
  transmute(
    name,
    obs_rate_nhb = sprintf("%0.1f (%0.1f)",
                           round(obs_rate_nhb, 1),
                           round(obs_se_nhb, 1)),
    obs_rate_nhw = sprintf("%0.1f (%0.1f)",
                           round(obs_rate_nhw, 1),
                           round(obs_se_nhw, 1)),
    obs_rd = sprintf(
      "%0.1f (%0.1f, %0.1f)",
      round(obs_rd, 1),
      round(obs_rd_lower, 1),
      round(obs_rd_upper, 1)
    ),
    obs_rr = sprintf(
      "%0.2f (%0.2f, %0.2f)",
      round(obs_rr, 2),
      round(obs_rr_lower, 2),
      round(obs_rr_upper, 2)
    )
  ) %>%
  kable(
    booktabs = TRUE,
    col.names = c(
      "Area",
      "Non-Hispanic Black",
      "Non-Hispanic White",
      "Rate difference (95% CI)",
      "Rate ratio (95% CI)"
    )
  ) %>%
  kable_classic() %>%
  add_header_above(c("",
                     "Age-standardized rates (SE)" = 2,
                     "",
                     ""))
```

## Table S12. 2019 rates, rate differences, and rate ratios by city/county
```{r}
wide_df %>%
  filter(opioid_type == "opioid", nchar(abbrev) > 2) %>%
  arrange(abbrev) %>%
  transmute(
    abbrev,
    obs_rate_nhb = sprintf("%0.1f (%0.1f)",
                           round(obs_rate_nhb, 1),
                           round(obs_se_nhb, 1)),
    obs_rate_nhw = sprintf("%0.1f (%0.1f)",
                           round(obs_rate_nhw, 1),
                           round(obs_se_nhw, 1)),
    obs_rd = sprintf(
      "%0.1f (%0.1f, %0.1f)",
      round(obs_rd, 1),
      round(obs_rd_lower, 1),
      round(obs_rd_upper, 1)
    ),
    obs_rr = sprintf(
      "%0.2f (%0.2f, %0.2f)",
      round(obs_rr, 2),
      round(obs_rr_lower, 2),
      round(obs_rr_upper, 2)
    )
  ) %>%
  kable(
    booktabs = TRUE,
    col.names = c(
      "Area",
      "Non-Hispanic Black",
      "Non-Hispanic White",
      "Rate difference (95% CI)",
      "Rate ratio (95% CI)"
    )
  ) %>%
  kable_classic() %>%
  add_header_above(c("",
                     "Age-standardized rates (SE)" = 2,
                     "",
                     ""))
```
